---
title: "HW 1 - SS 9159"
author: "Ravin Lathigra - 250620601"
date: "September 17, 2018"
output:pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=200)
```

```{r libraries, include=FALSE}
library(corrplot)    #Visualize Correlation withii
library(kableExtra)
library(tidyverse) 
library(dplyr)       #Manipulate data.frames
library(ggplot2)
library(ggbiplot)
library(formatR)
library(e1071)
library(VIM)
library(latex2exp)
```

*import data*

```{r import, eval = TRUE, echo = FALSE}
hw1_data <- read_csv("https://raw.githubusercontent.com/hgweon2/mda9159a/master/hw1_data1.csv")


knitr:: kable(hw1_data, caption = "HW1 Data", align = rep("c", ncol(hw1_data))) %>%
  kableExtra:: kable_styling() %>%
  scroll_box(height = "300px", width = "100%")

```

## Question 1

###A - x1 observations greater than 6

```{r 1a, eval= TRUE, echo = true}

one_a <- hw1_data %>% 
  filter(x1>6) %>%
  dplyr::summarise('Number of Observations where x1 > 6' = n())


kable(one_a, caption = "Question 1a", align = rep("c", ncol(one_a))) %>%
  kable_styling()

```

\n

The table above illustrates that within the data, there are 26 observations where x1 >6.

\n


###B - x1 observations greater than 6 and x2 = H

```{r 1b, eval= TRUE, echo = true}

one_b <- hw1_data %>%
  filter(x1>6) %>%
  filter(x2=="H") %>%
  dplyr::summarise('Number of observations such that x1 > 6 and X2 = H'= n())


kable(one_b, 
      caption = "Question 1b",
      align = rep("c", ncol(one_b))) %>%
  kable_styling() 
```

\n
\n

The table above illustrates that within the data, there are 23 observations where x1 >6 and x2 = H.

\n
\n


###C - Summary statistics of x1 conditioned on x2 = H

```{r 1c, eval= TRUE, echo = true}

A <- hw1_data %>%
  filter(x2=="H") %>%
  dplyr::summarise('Mean' = round(mean(x1),2), 'Median' = round(median(x1),2), 'Std. Deviation' = round(sd(x1),2))

kable(A, 
      caption = "Summary Statistics for x1 given x2 = H",
      align = rep("c", ncol(A))) %>%
  kable_styling() 
```



\n
\n

###D - t-Test of the true mean of x1

$$H_{0}: \mu  = 4$$
$$H_{a}: \mu  \ne 4$$
\n

$$\alpha = 0.05$$

\n


```{r 1d, eval= TRUE, echo = true}

x1 <- hw1_data$x1                                                  

x1_t_test <- t.test(x = x1,                                     #Perform two sided t-test.
                    alternative = "two.sided",
                    mu = 4,
                    conf.level = 0.95)
 
cv <- x1_t_test$statistic                                       #Store critical value from t-test.

x <- rt(10000, 99)                                              #generate samples from a t-distribution for visualization
  
y <- data.frame(t= x, t_density = dt(x, 99))                    


y<- y %>%                                                       #Add identifier to data to identify if observations are more extreme than critical values
  mutate(cutoffs = ifelse(abs(t)>cv,"True","False")) 

y_false <- y%>%
  filter(abs(t)<cv)
y_true_less<- y %>%
  filter(-t>=cv)
y_true_greater<- y %>%
  filter(t>=cv)

ggplot() +                                                      #Plot t-distributions
  geom_ribbon(data = y,aes(x = t, 
                  y = t_density, ymin = 0, 
                  ymax = t_density),
                  fill = "grey",
                  show.legend = FALSE) +
  geom_line(data = y,aes(x = t, 
                         y = t_density)) +
  geom_segment(data = y,aes(x =cv,                              #show critical value on plot
                            y = 0, 
                            xend = cv, 
                            yend = dt(-cv,99)))  +
  geom_segment(data = y,aes(x =-cv,                             #show critical values on plot
                            y = 0, 
                            xend = -cv, 
                            yend = dt(cv,99))) +
  geom_segment(data = y,aes(x = min(x),                         #frame the plot with horizontal line
                            y = 0, 
                            xend = max(x), 
                            yend = 0 )) +
  
  geom_segment(data = y,aes(x = min(x),                         #Add arrow to show area less than -cv
                            y = .1, 
                            xend = -cv, 
                            yend = .1 ),
               arrow = arrow(length = unit(0.5, "cm"), 
                             ends= "first")) +
  geom_segment(data = y,aes(x = cv,                             #Add arrow to show area less than -cv 
                            y = .1, 
                            xend = max(x), 
                            yend = .1 ),
               arrow = arrow(length = unit(0.5, "cm"),
                            ends= "last")) +
 
  
  geom_area(data = y_true_less,                                 #Fill in area under the t-distribution less than critical value
            aes(x=y_true_less$t, 
                y=y_true_less$t_density), 
            fill="blue", alpha= .3) +
  geom_area(data = y_true_greater,                              #Fill in area under the t-distribution greater than critical value
            aes(x=y_true_greater$t, 
                y=y_true_greater$t_density), 
            fill="blue", alpha= .3) +
  
  xlab("x") +
  ylab("Density") +
  ggtitle("PDF of t-distribution") +
  labs(subtitle = "Degrees of Freedom: 99") +
  
  annotate("text", x =(min(x)+-cv)/2, y = .125, label = "Pr(x<-cv) = 2.5%") +
  annotate("text", x =(cv+max(x))/2, y = .125, label = "Pr(x>cv) = 2.5%") +
  annotate("text", x = cv, y = -0.01, label = "cv") +
  annotate("text", x = -cv, y = -0.01, label = "-cv") +
  theme_bw()

print(x1_t_test)

```
 
 \n
 
 Using a two-sided t test, the  p-value was larger than the significance level which indicates that there is no significant evidence against the Null hypothesis, therefore we fail to reject the Null.
 
 \n
 \n
 
###E - T-Test of the true mean of x1 where x2 = H

$$H_{0}: \mu  \gt 4$$

$$H_{a}: \mu  \le 4$$
\n

$$\alpha = 0.05$$

 
 
 
```{r 1e, eval= TRUE, echo = true}



one_e <- data.frame(hw1_data %>%                                #Gather data needed for t-test.
  filter(x2 == "H") %>%
  select(x1))
  

  x1_t_test <- t.test(x = one_e,                                #Perform t-test
                      alternative = "less",
                      mu = 4,
                      conf.level = 0.95)

cv <- x1_t_test$statistic

x <- rt(10000, 99)

y<- y %>%                                                        #Add identifier to data to show when values exceeed the critical values
  mutate(cutoffs = ifelse(abs(t)>cv,"True","False")) 

y_false <- y%>%
  filter(abs(t)<cv)
y_true_less<- y %>%
  filter(-t>=cv)
y_true_greater<- y %>%
  filter(t>=cv)

ggplot() +                                                       #plot t-distribution and add labels
  geom_ribbon(data = y,aes(x = t, 
                  y = t_density, ymin = 0, 
                  ymax = t_density),
                  fill = "grey",
                  show.legend = FALSE) +
  geom_line(data = y,aes(x = t, y = t_density)) +
  geom_segment(data = y,aes(x =cv, y = 0, xend = cv, yend = dt(-cv,99)))  +
  geom_segment(data = y,aes(x =-cv, y = 0, xend = -cv, yend = dt(cv,99))) +
  geom_segment(data = y,aes(x = min(x), y = 0, xend = max(x), yend = 0 )) +
  
  geom_area(data = y_true_less,aes(x=y_true_less$t, y=y_true_less$t_density), fill="blue", alpha= .3) +
  geom_area(data = y_true_greater,aes(x=y_true_greater$t, y=y_true_greater$t_density), fill="blue", alpha= .3) +
  
  xlab("x") +
  ylab("Density") +
  ggtitle("PDF of t-distribution") +
  labs(subtitle = "Degrees of Freedom: 99") +

  annotate("text", x = cv, y = -0.01, label = "cv") +
  annotate("text", x = -cv, y = -0.01, label = "-cv") +
  theme_bw()

  print(x1_t_test)

```
\n

Considering a one-sided t-test, the p-value exceeds the significance level of 5% which suggest that we fail to reject the Null hypothesis that the mean of x1 given x2 = H is greater than 4.

\n
\n

## Question 2

```{r 2 setup, eval= TRUE, results = "hide"}

set.seed(50)
idx <- sample(nrow(cars), 40, replace = FALSE)
cars2 <- cars[idx,]

```

\n

###A - Relationship between Speed and Distance

```{r 2a, eval= TRUE, echo = true}

spd_dist_plot <- ggplot(cars2)+         
                    geom_point(aes(x = speed, y = dist)) +     #Scatterplot of Cars Data
                    geom_smooth(aes(x = speed, y = dist)) +    #Apply 95% C.I using lm
                    ggtitle("Cars2 - Distance Vs Speed") +    
                    xlab("Speed") +
                    ylab("Distance") +
                    theme_bw()

print(spd_dist_plot)
```
\n

From the scatter plot "Cars2 - Distance Vs Speed" we can see that there is a positive correlation between speed and distance.  Superimposing a smooth plot onto the existing plot with a 95% confidence interval aids in ilustrating there is a relationship between speed and distance and that modelling this relationship with a linear model may be appropriate.

\n
\n

###B - Least Squares Estimates

```{r 2b, eval= TRUE, echo = true}

two_b_lm <- lm(dist ~ speed, data= cars2)               #Create Linear Model

ls_sigma <- summary(two_b_lm)$sigma^2                   #Extract LS Estimate for Sigma
ls_betas <- summary(two_b_lm)$coefficients[,1]          #Extract LS Estimate for Beta0, Beta1

beta <- intToUtf8(946)                                  #Convert Unicode for greek beta into symbol
sigma_sq <- intToUtf8(963)                              #Convert Unicode for greek sigma into symbol
 
ls_estimates <- data.frame(                             #Create Data.Frame for LS estimates
                  cbind(ls_betas[1],
                        ls_betas[2],
                        ls_sigma)
                        )     

colnames(ls_estimates) = c(paste0(beta,c(0:1)),         #Apply greek symbols to column Headers
                           paste0(sigma_sq,"^2"))
rownames(ls_estimates) = "Estimates"                    #Rename row

kable(ls_estimates,                                     #Create table smmarizing LS esimates for output
      caption = "Least Squares Estimates: Distance ~ Speed",
      align = rep("c", ncol(ls_estimates))) %>%
  kable_styling() 


```

\n
\n

###C - Calculating 4th, 7th and 10th residuals

```{r 2c, eval= TRUE, echo = true}

resid <-two_b_lm$residuals                              #Extract residuals from linear model


cars2 <- cars2 %>%                                      #Append Residuals to Cars2 data
  mutate(residual = resid)


ggplot(cars2) +
  geom_point(aes(x = cars2$speed, two_b_lm$residuals))+ #Plot Speed vs residuals
  
  geom_segment(aes(x = cars2$speed[4],                  #Draw line segment showing 4th residual
                   y = 0, 
                   xend = cars2$speed[4], 
                   yend = cars2$residual[4]), 
               colour = "darkblue") +
  
  geom_segment(aes(x = cars2$speed[7],                  #Draw line segment showing 7th residual 
                   y = 0, 
                   xend = cars2$speed[7], 
                   yend = cars2$residual[7]), 
               colour = "darkblue") +
  
  geom_segment(aes(x = cars2$speed[10],                 #Draw line segment showing 10th residual 
                   y = 0, 
                   xend = cars2$speed[10], 
                   yend = cars2$residual[10]), 
               colour = "darkblue") +
  
  geom_point(data = cars2[c(4,7,10),],                  #Recolour 4th, 7th, and 10th residuals
             aes(x = cars2$speed[c(4,7,10)], 
                 y= cars2$residual[c(4,7,10)]), 
                 colour = "red", size = 2, 
             show.legend = FALSE) +
  
  geom_text(aes(x = cars2$speed[c(4)],                  #Label 4th residuals
                y= cars2$residual[c(4)],
                label = "4th Residual",
                alpha = 1), 
            show.legend = FALSE , 
            size = 3,
            nudge_y = -2,  
            colour = "Blue") +
  
  geom_text(aes(x = cars2$speed[c(7)],                  #Label 7th residuals
                y= cars2$residual[c(7)],
                label = "7th Residual",
                alpha = 1), 
            show.legend = FALSE , 
            size = 3,
            nudge_y = 2,
            colour = "Blue") +
  
  geom_text(aes(x = cars2$speed[c(10)],                 #Label 10th residuals 
                y= cars2$residual[c(10)],
                label = "10th Residual",
                alpha = 1), 
            show.legend = FALSE , 
            size = 3,
            nudge_y = 2,  
            colour = "Blue") +

  geom_hline(yintercept = 0, colour= "darkblue") +      #Add horizontal line at y=0
  
  ggtitle("Cars2 - Speed vs Residuals") +
  
  labs(x  = "Speed", 
       y  = "Residual") +
  
  theme_bw() 


cars_residuals <- data.frame(cars2$residual[c(4,7,10)])

rownames(cars_residuals) = c("4th Residual", 
                             "7th Residual",
                             "10th Residual")
colnames(cars_residuals) = c("Residual")


kable(cars_residuals,                                     #Create table smmarizing LS esimates for output
      caption = "4th, 7th and 10th Residuals",
      align = rep("c", ncol(cars_residuals))) %>%
  kable_styling() 


```


\n
\n

###D - Indicating residuals >20.
\n

```{r 2d part 1, eval= TRUE, echo = true}



cars_w_residual_size <- cars2 %>%                          #add an identifier to Cars2 data inidcating the size of residuals i.e >20 or <20.
                mutate(ResidualSize = ifelse(abs(residual)>20,">20","<=20"))

cars2_plot <-                                              #plot data, applying breaks to allow aesthetics to be applied based on the size of residuals
  ggplot(cars_w_residual_size,
         aes(x = speed, 
             y = dist, 
             group = ResidualSize, 
             size = ResidualSize,
             shape = ResidualSize, 
             colour = ResidualSize))+          
  geom_point() +
  scale_colour_manual(breaks = c("<=20",">20"),            
                      values = c("red", "blue"))+          #Apply different colours based on residual size
  scale_size_manual(breaks = c("<=20",">20"),
                    values = c(3, 3.5))+                   #Scale the plotted points

  ggtitle("Cars2 - Distance Vs Speed") +                   #Change Plot titles & theme
  xlab("Speed") +
  ylab("Distance") +
  theme_bw()

cars2_plot

```

\n 
The above plot shows speed vs distance for the Cars2 data.  Observations that had residuals larger than 20 were highlighted in blue.

\n

We can better understand the residuals of the regression line better if we plot the observations against their residuals.  The following plot shows the relationship between residuals and speed.

```{r 2d part 2, eval= TRUE, echo = true}


large_resid <- cars2 %>%                                   #filter to Cars2 data by size of residuals i.e >20.
                filter( abs(residual)>20)
  
cars2_plot_residual <- ggplot(cars2) +                     #Create residual plot of data
  
  geom_point(aes(x = cars2$speed, 
                 y=two_b_lm$residuals))+
  
  geom_point(data = large_resid,
             aes(x = speed, 
                 y= residual), 
                 colour = "red", 
                 shape = 3,
                 size = 3,
             show.legend = TRUE) +

  geom_hline(yintercept = 0,                               #Insert horizontal line at y=0
             colour= "darkblue") + 
  
  geom_hline(yintercept = 20,                              #Insert horizontal line at y=20
             linetype = "dotted", 
             colour = "darkblue") +
  
  geom_hline(yintercept = -20,                             #Insert horizontal line at y=-20
             linetype = "dotted", 
             colour = "darkblue") +

  ggtitle("Cars2 - Speed vs Residuals") +                  #Add titles and themes
  
  labs(x  = "Speed", 
       y  = "Residual") +
  
  theme_bw() 

cars2_plot_residual

```

It is worth noting that apart from the 7 observations that had large residuals i.e above 20, the remaining 82.5% of the data was within 20 and showed no general trend amongst the residuals.

\n

###E - Sum of Residuals
\n

```{r 2e, eval= TRUE, echo = true}



sum_of_residuals <- sum(cars2$residual)                    #Calculate sum of residuals

cat("Sum of residuals =", round(sum_of_residuals,5))
```

\n
\n

###F - Plotting fitted Line & Predicting values
\n


```{r 2f, eval= TRUE, echo = TRUE}


beta0 <- two_b_lm$coefficients[1]                          #Extract coefficients from lm.   
beta1 <- two_b_lm$coefficients[2]

cars2_plot +                                                
  
  geom_line(aes(x = speed, y = beta0 + beta1*speed),       #Add the fitted model to the existing cars2_plot
            linetype = "solid", 
            colour = "darkblue", 
            size = 1, 
            show.legend = FALSE)+
  
  scale_fill_discrete("")+                                 #Remove legend title
  
  geom_point(aes(x = 17,                                   #Add dotted lines showing predicted value when speed =17
                 y = predict(two_b_lm, 
                             newdata = data.frame(speed = 17))),
             colour = "green3", size = 3, shape = 17) +
  
  geom_segment(aes(x = 17,                                 #Add dotted lines showing predicted value when speed =17
                   xend = 17, 
                   y = -Inf, 
                   yend = predict(two_b_lm, newdata = data.frame(speed = 17))), 
               colour= "darkgrey", 
               linetype = "dotted", 
               size = .75) +
  
  geom_segment(aes(x = -Inf,                               #Add dotted lines showing predicted value when speed =17
                   xend = 17, 
                   y = predict(two_b_lm, newdata = data.frame(speed = 17)), 
                   yend = predict(two_b_lm, newdata = data.frame(speed = 17))), 
               colour= "darkgrey", 
               linetype = "dotted", 
               size = .75) +
  
  annotate("text", x = 17,                                 #Add text to plot
           y = -40, 
           label = "Speed = 17", 
           size = 3)  +
  
  annotate("text",                                         #Add text to plot
           x = min(cars2$speed)+1, 
           y = predict(two_b_lm, 
                       newdata = data.frame(speed = 17)), 
           label = "Distance = 48.8", size = 3)  +

  ggtitle("Cars2 - Distance Vs Speed with Fitted Model")   #Add title


fit_model <- predict(two_b_lm, newdata = data.frame(speed = 17)) #Predict using the fittded model distance to stop given a speed of 17 mph.


cat("Considering the fitted model, the predicted distance taken to stop when the speed of the car is 17 mph is", round(fit_model,2))


```



\n

The LS estimate for the the regression line that models distance as a function of speed for the car2 dataset is:

$$\hat{Y} = \hat{\beta_{0}} + \hat{\beta_{1}}X$$ where;

\n

$$\hat{\beta_{0}} = -17.24$$
$$\hat{\beta_{1}} = 3.88$$
\n
\n

###G - Goodness of fit

\n

```{r 2g, eval = TRUE, echo=TRUE}

SSE<- sum((cars2$residual)^2)

SST <- sum((cars2[,c("dist")]-mean(cars2$dist))^2)

r_2 <- 1-SSE/SST

paste("Proportion of Variance explained by the regression model",round(r_2,2), sep = "-> ")

```
\n

###H - Extrapolating Data

\n

```{r 2h, eval = TRUE, echo=TRUE}

cars2_speed_summary <- rbind(data.frame(min(cars2$speed),     #Generate key summary statistics of the domain of the cars2 speed data.
                                        mean(cars2$speed), 
                                        max(cars2$speed),
                                        sd(cars2$speed)))


fit_model_100 <- predict(two_b_lm,                             #USing the fitted model, predict the outcome considering a speed of 100mph.
                         newdata = data.frame(speed = 100))


rownames(cars2_speed_summary) = c("Statistics")
colnames(cars2_speed_summary) = c("Min", "Mean", "Max", "Standard Dev.")


paste("Considering the regression model for distance, at 100km/h we predict that the response i.e the distance to be: ", round(fit_model_100,2), "Km", sep = " ")


kable(cars2_speed_summary, caption = "Domain statistics of Speed for Cars2 data",
      align = rep("c", ncol(cars2_speed_summary))) %>%
  kable_styling(position = "center") 



```
\n 

The summary provied above displays the domain of the sample data for the Cars2 dataset.  Generally, infering data from the regression line is suitable however, exprapolating data from the regression line can lead to poor predictions.  In this case, making predictions for the distance at 100mph is an unreliable prediction as the support for the model has a maximum value of 25 mph, and a minimum of 4mph.  100mph is several standard deviations from the mean and outside the max speed in the training data.  Therefore, making decisions, or drawing conclusions from this data is irreponsible.

\n
\n

###I - Relationship between x and Y - Hypothesis Testing

\n


```{r 2i, eval = TRUE, echo=TRUE}

test_b1 <- summary(two_b_lm)

print(test_b1)


```

\n

Considering the following Null and Alternative Hypothesis:

$$H_{0} : \beta_{1} = 0 $$;
$$H_{A} : \beta_{1} \ne 0 $$

and significance level:

$$\alpha = 0.05$$

The above summary shows that there is strong evidence against the Null hypothesis, therefore we reject the Null hypothesis.

\n
\n


###J - Hypothesis Testing - One sided

\n


```{r 2j, eval = TRUE, echo=TRUE}


beta1_true <- 4

se_beta1 <- summary(two_b_lm)$coefficients[2,2]

t_stat <- (beta1-beta1_true)/(se_beta1)

pvalue <- pt(t_stat, 38)  


t_test_df <- data.frame('Beta 1 hat' = beta1, 'Beta 1 null' = beta1_true, 'Standard Error' = se_beta1, 't-stat' = t_stat, 'Pr(>|t|)' = pvalue)

colnames(t_test_df) = c("Beta Hat", "Beta-Null", "Standard Error", "t Stat", "Pr(>|t|)")
rownames(t_test_df) = c("Key Values")

kable(t_test_df,
      caption = "t-Test Statistics",
      align = rep("c", ncol(t_test_df))) %>%
  kable_styling()
```
\n

Considering the following Null and Alternative Hypothesis:

$$H_{0} : \beta_{1} = 4 $$;
$$H_{A} : \beta_{1} \le 4 $$

and significance level:

$$\alpha = 0.05$$
\n

The above table, shows the p-value is significantly greater than significance level which indicates that there is no evidence against the Null Hypothesis.   Therefore, we fail to reject the Null.  
